### 2016-02-15 acknowledgments.tex

\section{Acknowledgments}
We would like to thank Richardson lab members Dan Keedy and Bryan Arendall for setting up the chain-level Top8000, Jeff Headd for moving the MolProbity utilities to use the \textit{PHENIX} cctbx toolbox, and Michael Prisant for help with GitHub. Also, thanks to \textit{PHENIX} Project members Nigel Moriarty and Nat Echols for help with integration into \textit{PHENIX} automation and GUI. This work was supported by NIH grants R01-GM073919 and ProjectIV of P01-GM063210, and by an NSF Graduate Research Fellowship to BJH.