### 2016-02-15 title_page.tex

\title{MolProbity's Ultimate Rotamer-Library Distributions for Model Validation}
\author{Bradley J. Hintze}
\author{Steven M. Lewis}
\author{Jane S. Richardson}
\author{David C. Richardson}
\affil{Department of Biochemistry, Duke University}
\maketitle

\noindent
Short title: MolProbity's Ultimate Rotamer-Library
\newline
\newline
\noindent
Keywords: sidechain rotamer library, rare sidechain conformations, structural bioinformatics, structure validation, Phenix, high-quality dataset, protein conformation
\newline
\newline
\noindent
All work performed at Duke University, Durham, North Carolina.
\newline
\newline
\noindent
Correspondence to: David C. Richardson, 132 Nanaline Duke Building, Duke University, Durham NC 27710. E-mail: dcrjsr@kinemage.biochem.duke.edu