### 2016-02-15 supplemental/supp.tex
added the rest of the contour compare figures to supplement

% Bradley Hintze -- June 2013
\documentclass{article}

%\usepackage{rough_draft}
\usepackage{gensymb,wrapfig,booktabs,amssymb}
\bibliographystyle{unsrt}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{fixltx2e}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[bottom=1in, top=1in, left=0.75in, right=0.75in]{geometry}
\usepackage{pdflscape}
\usepackage{supertabular}
\usepackage{xcolor,colortbl}
\graphicspath{{figures/}}
%\linespread{2}
%-----------------------------------------------------------------------------%
% HYPERREF: plain black hypertext references for ref's and cite's.
%-----------------------------------------------------------------------------%
\usepackage[pdftex, pdfusetitle, plainpages=false, 
                                letterpaper, bookmarks, bookmarksnumbered,
                                colorlinks, linkcolor=black, citecolor=black,
                 filecolor=black, urlcolor=black]
                                {hyperref}

\definecolor{Gray}{gray}{0.85}
\definecolor{LightCyan}{rgb}{0.88,1,1}

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}

\begin{document}

\title{Rough Draft of Supplemental Material}
\author{Bradley Hintze}
\date{January 2015}
\maketitle
\tableofcontents

\newpage
\input{residuecounts}
\newpage
\section{Top500 vs. Top8000 Reference Contours}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.75\textwidth]{contour_compare_sup}
  \caption{Areas in orange (from Top500 data) and in blue (from Top8000) fill
      the allowed regions for AsN, His, Leu, and Phe/Tyr. Phe and Tyr were done
      together as no significant difference was detected between their distributions.
      The extensive areas in green are where the two systems both declare allowed 
      conformations.}
    \label{fig:outcounts}
\end{figure}
\newpage
\input{out_8000_500}
\newpage
\input{frequency}
\newpage
\input{chis_and_angles}
\newpage
%\input{pdbchains}

\end{document}