### 2016-02-22 supplemental/supp.tex
Fixed Supplement table and figure numbers and how they are referred to in the main text

\documentclass{article}

\usepackage{gensymb,wrapfig,booktabs,amssymb}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{fixltx2e}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[bottom=1in, top=1in, left=0.75in, right=0.75in]{geometry}
\usepackage{pdflscape}
\usepackage{supertabular}
\usepackage{xcolor,colortbl}
\graphicspath{{figures/}}
%\linespread{2}
%-----------------------------------------------------------------------------%
% HYPERREF: plain black hypertext references for ref's and cite's.
%-----------------------------------------------------------------------------%
\usepackage[pdftex, pdfusetitle, plainpages=false, 
                                letterpaper, bookmarks, bookmarksnumbered,
                                colorlinks, linkcolor=black, citecolor=black,
                 filecolor=black, urlcolor=black]
                                {hyperref}
\usepackage{authblk}

\definecolor{Gray}{gray}{0.85}
\definecolor{LightCyan}{rgb}{0.88,1,1}

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thesection}{S\arabic{section}}
\begin{document}

\title{Supplemental Material for MolProbity's Ultimate Rotamer-Library Distributions for Model Validation}
\author{Bradley J. Hintze}
\author{Steven M. Lewis}
\author{Jane S. Richardson}
\author{David C. Richardson}
\affil{Department of Biochemistry, Duke University}
\date{}
\maketitle
\listoffigures
\listoftables
\newpage

\section*{Top500 vs. Top8000 Reference Contours}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.75\textwidth]{contour_compare_sup}
  \caption[Top500 vs Top8000 Contours for Residues with 2 $\chi$s ]
      {Areas in orange (from Top500 data) and in blue (from Top8000) fill
      the allowed regions for AsN, His, Leu, and Phe/Tyr. Phe and Tyr were done
      together as no significant difference was detected between their distributions.
      The extensive areas in green are where the two systems both declare allowed 
      conformations. See Figure 2 for Asp, Trp, and Ile}
    \label{fig:outcounts}
\end{figure}
\newpage

\input{residuecounts}
\newpage
\input{out_8000_500}
\newpage
\input{frequency}
\newpage
\input{chis_and_angles}
\newpage

\end{document}